### 2018-04-11 paper.Rmd
builds, reorg images
spell check plus abstract

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. However, traditional methods of inquiry into the data have often relied on a command line interface to extract developmental indices or custom parsers to extract data for computational modeling. Now, advances in open-source software allow for novel approaches in working with this dataset. We introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API).  Along with a versioned infrastructure that facilitates reproduction of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

```{r load_packages, echo = FALSE, warning=FALSE}
library(papaja)
library(tidyverse)
library(childesr)
library(directlabels)
library(koRpus)
library(png)
library(here)
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
```

@@

We begin by describing the architecture of childes-db and the web applications that we provide. We next describe the `childesr` API, which allows for programmatic access to the data. We then present two worked examples of specific uses of the system using both the web app and the API. The first examines the frequency of color words in children's input; the second quantifies the growth of lexical diversity across genders. We end by presenting some suggestions for the use of the childes-db web frontend in classroom teaching.

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
Here we introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific hsitorical versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? -->
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions-->
<!-- SM: can people get to these sql files currently on the website?-->

@@

For this analysis, we will define our words of interest as the basic color words of English (except for gray, which children hear very rarely). We store these in the `colors` variable, and then use the `get_types` function from `childesr` to get the type frequency of each of these words in all of the corpora in CHILDES. For demonstration, we look only at the types produced by the speakers in each corpus oftagged as Mother and Father. We also restrict ourselves to children from 1-5 years old (12 to 60 months), and look only the American English corpora.

@@

To normalize correctly (i.e. to ask what proprotion of the input children hear consists of these color words), we need to know how many total words these children hear from their parents in these transcripts. To do this, we use the `get_speaker_statistics` function, which will return a total number of tokens (`num_tokens`) for each of these speakers.

@@

mod <- lme4::lmer(measure ~ target_child_sex * poly(age_years, 2) - 1 + 
                    (1 | target_child_id) +
                    (target_child_sex + age_years | language), 
  data = data)

@@

png::readPNG(here::here("imgs/in_class_teach_viz.png")) %>% grid::grid.raster()
```

@@

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies presented above -- color words or gender. Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` starter code to provide before asking students to generate their own plots and writeups. Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic], building in pre-written blocks of executable code that demonstrate key programming concepts for students and including incomplete or blank code blocks that students can use to write their own code.

>>>>

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. However, current methods of using CHILDES data rely on a command line interface to extract developmental indices or custom parsers to extract data for larger-scale analyses. We introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API). Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

```{r load_packages, echo = FALSE, warning=FALSE}
library(papaja)
library(tidyverse)
library(childesr)
library(directlabels)
library(koRpus)
library(png)
library(here)
library(lme4)
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)
```

@@

We begin by describing the architecture of childes-db and the web applications that we provide. We next describe the `childesr` API, which allows for programmatic access to the data. We then present two worked examples of specific uses of the system using both the web app and the API. The first examines the frequency of color words in children's input; the second quantifies the growth of lexical diversity across genders. We end by presenting some suggestions for the use of the childes-db web front-end in classroom teaching.

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
Here we introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? -->
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions-->
<!-- SM: can people get to these sql files currently on the website?-->

@@

For this analysis, we will define our words of interest as the basic color words of English (except for gray, which children hear very rarely). We store these in the `colors` variable, and then use the `get_types` function from `childesr` to get the type frequency of each of these words in all of the corpora in CHILDES. For demonstration, we look only at the types produced by the speakers in each corpus tagged as Mother and Father. We also restrict ourselves to children from 1-5 years old (12 to 60 months), and look only the American English corpora.

@@

To normalize correctly (i.e. to ask what proportion of the input children hear consists of these color words), we need to know how many total words these children hear from their parents in these transcripts. To do this, we use the `get_speaker_statistics` function, which will return a total number of tokens (`num_tokens`) for each of these speakers.

@@

mod <- lmer(measure ~ target_child_sex * poly(age_years, 2) - 1 + 
                    (1 | target_child_id) +
                    (target_child_sex + age_years | language), 
  data = data)

@@

png::readPNG(here::here("images/in_class_teach_viz.png")) %>% grid::grid.raster()
```

@@

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies presented above -- color words or gender. Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` starter code to provide before asking students to generate their own plots and write-ups. Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic], building in pre-written blocks of executable code that demonstrate key programming concepts for students and including incomplete or blank code blocks that students can use to write their own code.
